*Supplementary Material 1*

### 2022-02-24 ms/manuscript.Rmd
Update MS

We rooted the ML tree on bryophytes and estimated divergence dates using penalized likelihood with treePL [@Smith2012].
We selected `r n_calibrations_ferns` fern fossils to use as constraints from a curated list of `r n_fern_fossils_total` fern fossil taxa (`r s_table("fossils")`, Appendix S2).
In the case of redundant fossils (those assignable to the same node in the phylogeny; `r eg`, stems of families that are sister to each other, such as Dipteridaceae and Matoniaceae) we selected the fossil with the oldest age (`r ie`, the upper limit of the oldest stratigraphic period assigned to the fossil). 
We assigned fossils to lineages only after consulting the original publication (as opposed to simply relying on the fossil name) so we could reassess the identification relative to changes in taxonomy and hypotheses of phylogenetic relationships.
Taxonomic concepts may vary between the original publication and currently applied taxonomy (`r eg`, Dicksoniaceae *sensu lato* including other tree fern families used in description of the fossil but Dicksoniaceae *sensu stricto* used currently) and hypotheses of phylogenetic relationships among extant species may change (`r eg`, *Dennstaedtia* used in description of a fossil but this genus now known to be polyphyletic); in both cases, the resulting node to be constrained for a given fossil could change.
We also applied two constraints outside of ferns, including the most recent common ancestor (MRCA) of euphyllophytes (`r fossils$Euphyllophytes$age` Ma) and MRCA of tracheophytes (`r fossils$Tracheophytes$age` Ma), and fixed the root age (land plants) at 475 Ma as in @Testo2016a and @Qi2018.
All constraints other than the root are minimum ages (`r s_figure("fossil_map")`).

>>>>

We rooted the ML tree on bryophytes and estimated divergence dates using penalized likelihood with treePL [@Smith2012].
We selected `r n_calibrations_ferns` fern fossils to use as constraints from a newly curated database of `r n_fern_fossils_total` fern fossil taxa (`r s_table("fossils")`, Appendix S2).
The full database is available at https://fernphy.github.com/FernCal.
In the case of redundant fossils (those assignable to the same node in the phylogeny; `r eg`, stems of families that are sister to each other, such as Dipteridaceae and Matoniaceae) we selected the fossil with the oldest age (`r ie`, the upper limit of the oldest stratigraphic period assigned to the fossil). 
We assigned fossils to lineages only after consulting the original publication (as opposed to simply relying on the fossil name) so we could reassess the identification relative to changes in taxonomy and hypotheses of phylogenetic relationships.
Taxonomic concepts may vary between the original publication and currently applied taxonomy (`r eg`, Dicksoniaceae *sensu lato* including other tree fern families used in description of the fossil but Dicksoniaceae *sensu stricto* used currently) and hypotheses of phylogenetic relationships among extant species may change (`r eg`, *Dennstaedtia* used in description of a fossil but this genus now known to be polyphyletic); in both cases, the resulting node to be constrained for a given fossil could change.
We also applied two constraints outside of ferns, including the most recent common ancestor (MRCA) of euphyllophytes (`r fossils$Euphyllophytes$age` Ma) and MRCA of tracheophytes (`r fossils$Tracheophytes$age` Ma), and fixed the root age (land plants) at 475 Ma as in @Testo2016a and @Qi2018.
All constraints other than the root are minimum ages (`r s_figure("fossil_map")`).

### 2022-02-24 ms/si.Rmd
Update MS

>>>>

```{r fossil-map}
# Make tibble with node number and label for ggtree
node_labs_tbl <-
  fossil_calibrations_for_treepl %>%
  transmute(
    fossil_label = str_remove_all(min, "min = "),
    node = map2_dbl(taxon_1, taxon_2, ~ape::getMRCA(sanger_tree_rooted, c(.x, .y))),
    calibration_type = case_when(
      !is.na(min) & is.na(max) ~ "minimum",
      !is.na(min) & !is.na(max) ~ "fixed",
      TRUE ~ NA_character_
    )
  ) %>%
  assert(not_na, calibration_type)

# Make plot
fossil_map_plot <- 
  ggtree(
    sanger_tree_rooted,
    branch.length = "none", size = 0.1) %<+% node_labs_tbl +
  # Add node points
  geom_nodepoint(
    aes(color = calibration_type), size = 1) +
  scale_color_manual(
    "Calibration type",
    values = c(
      minimum = palette_okabe_ito("vermillion"), 
      fixed = palette_okabe_ito("skyblue")),
    na.value = "transparent"
  ) +
  # Add node labels
  geom_text(aes(label = fossil_label), hjust = 0, size = 1.0, color = "grey20") +
  # Add tip labels
  geom_tiplab(aes(label = label), size = 0.5) +
  theme(legend.position = "bottom")

# Save plot
ggsave(
  plot = fossil_map_plot,
  file = (result_file(s_figure("fossil_map"), "pdf")),
  height = 60, width = 8, units = "in",
  limitsize = FALSE)
```

**`r s_figure("fossil_map")`**.
Maximum likelihood fern phylongey with node labels indicating placement of fossil constraints (`r result_file(s_figure("fossil_map"), "pdf") %>% fs::path_file()`).
Number for each constraint is age in Ma.
